---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r global_options , include=FALSE}
knitr::opts_chunk$set(fig.path='Figs/')
```

## Loading and preprocessing the data
##### 1. Load the necessary libraries
```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}

library(ggplot2)
library(scales)
library(Hmisc)

```
##### 2. Unzip and Loading file
```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}

if(!file.exists('activity.csv')){
  unzip('activity.zip')
}
data <- read.table('activity.csv', header = TRUE, sep = ',')

```
##### 2. Checking the data
```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}

summary(data)
str(data)
dim(data)

```
##### 2. Changing the type of date
```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}

data$date <- as.Date(data$date, "%Y-%m-%d")

```
## Mean total number of steps taken per day with histogram
##### 1. Histogram of total number of steps taken each day
```{r total_steps_per_day, echo=TRUE, results='hide', warning=FALSE, message=FALSE}

total.steps <- tapply(data$steps, data$date, FUN = sum, na.rm = TRUE)
qplot(total.steps, binwidth = 500, xlab ='total number of steps taken each day', ylab = 'Frequency with binwidth = 500')
mean(total.steps, na.rm = TRUE)
median(total.steps, na.rm = TRUE)

```
##### 2. Mean and median
```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}

Mean <- mean(total.steps, na.rm = TRUE)
Median <- median(total.steps, na.rm = TRUE)

```
* Mean     : `r Mean`
* Median   : `r Median`

----

## Average daily activity pattern
##### 1.Time series plot
```{r average_steps_taken, echo=TRUE, warning=FALSE, message=FALSE}

average <- aggregate(x = list(steps = data$steps), by = list(interval = data$interval),
                     FUN = mean, na.rm = TRUE)
ggplot(data = average, aes(x = interval, y = steps))+
  geom_line()+
  xlab('5 minutes interval')+
  ylab('average number of steps taken')

```

##### 2. 5-minute interval on average across all the days in the dataset that contains the maximum number of steps
```{r, echo=TRUE, warning=FALSE, message=FALSE}
mostSteps <- which.max(average$steps)
timeMostSteps <- average[mostSteps,][1,1]
answer <- gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2",timeMostSteps)
```
* Most Steps at: `r answer`

----

## Imputing missing values
##### 1. The total number of missing values
```{r, echo=TRUE, warning=FALSE, message=FALSE}

numMissingVal <- length(which(is.na(data$steps)))

```
* Number of missing values: `r numMissingVal`

##### 2. Replacing missing values with mean value of its 5-minute interval
```{r, echo=TRUE,warning=FALSE, message=FALSE}

fill.value <- function(steps,interval){
  filled <- NA
  if (!is.na(steps))
    filled<- c(steps)
  else
    filled <- (average[average$interval == interval, 'steps'])
  return (filled)
}
filled.data <- data
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)

```
##### 3. Using filled dataset to make histogram of total number of steps taken each day and calculate mean and median
```{r total_steps_per_day_after_filling_missing_values, echo=TRUE, warning=FALSE, message=FALSE}

total.steps <- tapply (filled.data$steps, filled.data$date, FUN = sum)
qplot(total.steps, binwidth = 500, xlab = 'total number of steps taken each day')

Meanafter <- mean(total.steps)
Medianafter <- median(total.steps)

```
* Mean (Imputed)    : `r Meanafter`
* Median (Imputed)  : `r Medianafter`
The values are higher then before with missing data because those missing values are replaced with mean value of `steps`.

## Are there differences in activity patterns between weekdays and weekends?
##### 1. Create new factor variable with two levels ??? ???weekday??? and ???weekend???
```{r, echo=TRUE,  warning=FALSE, message=FALSE}

weekday.or.weekend <- function(date){
  day <- weekdays(date)
  if (day %in% c('Monday','Tuesday','Wednesday','Thursday','Friday'))
    return('weekday')
  else if (day %in% c('Saturday','Sunday'))
    return('weekend')
  else
    return('invalid date')
}
filled.data$day <- sapply(filled.data$date, FUN = weekday.or.weekend)

```
##### 2. Panel plot containing a time series plot
```{r steps_Vs_interval_on_weekday_and_weekend, echo=TRUE,  warning=FALSE, message=FALSE}

average <- aggregate(steps ~ interval + day, data = filled.data ,mean)
ggplot(average,aes(interval, steps))+
  geom_line()+
  facet_grid(day ~ .)+
  xlab('5-minute interval')+
  ylab('Number of steps')


```

